Place of death in NRS COVID-19 deaths

### 2021-02-22 covid_nrs_place_of_death.Rmd
Added plots of covid cases per 100,000 on a map by HB/LA - they match the home deaths it seems!

Note: the cumulative case counts have already been calculated, we just need to subset them to match the weeks. This is not trivial - the death counts cover an entire week starting from the "w/c date"; the cumulative cases for that week could be taken either at the start, the end of the week, or the start of the following week even!
However, when calculating weekly positive cases we can simply take the maximum of that week's cumulative count - that will be the cumulative count at the end of the week!

@@

```{r}
areas_relatively_higher_number_home_deaths <- "Lighter areas represent areas with relatively higher number of home deaths in 2020 compared to historical mean"
home_deaths_ratio <- "Ratio computed from total home deaths in 2020 divided by mean annual home deaths 2015-2019"

@@

```{r, fig.height=12, fig.width=8}
# TODO: work out the size of these plots
figure_map_ratio_home_deaths_2020_to_historical_la <-
  total_deaths_la %>%
  filter(
    year == 2020 & place_of_death == "Home & other non-institution"
  ) %>%
  left_join(shapefile_la, by=c("ref_area"="LAD20NM")) %>%
  ggplot(aes(geometry=geometry, fill=ratio_annual_deaths_to_historical_mean), colour="grey10") +
  geom_sf(colour="grey10") +
  scale_fill_viridis_b(limits = c(minimum_ratio_between_hb_and_la_home_deaths, maximum_ratio_between_hb_and_la_home_deaths)) +
  theme_minimal() +
  theme(axis.text = element_blank()) +
  labs(x="", y="",
       title = "Ratio of home deaths in 2020 to historical mean 2015-2019, by Local Authority",
       subtitle = "A ratio of 1 means there was no increase; A ratio of 1.6 means there was a 60% increase of home deaths in 2020 over the mean 2015-2019.",
       caption = paste0(c(areas_relatively_higher_number_home_deaths, home_deaths_ratio, source_nrs_text), collapse="\n"),
       fill = "Ratio"
       ) +
  NULL

@@

```{r, fig.height=12, fig.width=8}
# TODO: work out the size of these plots
figure_map_ratio_home_deaths_2020_to_historical_hb <-
  total_deaths_hb %>%
  filter(
    year == 2020 & place_of_death == "Home & other non-institution"
  ) %>%
  left_join(shapefile_hb_uk_wide, by=c("ref_area"="name")) %>%
  ggplot(aes(geometry=geometry, fill=ratio_annual_deaths_to_historical_mean), colour="grey10") +
  geom_sf(colour="grey10") +
  scale_fill_viridis_b(limits = c(minimum_ratio_between_hb_and_la_home_deaths, maximum_ratio_between_hb_and_la_home_deaths)) +
  theme_minimal() +
  theme(axis.text = element_blank(), aspect.ratio = 1.4) +
  labs(x="", y="",
       title = "Ratio of home deaths in 2020 to historical mean 2015-2019, by Health Board",
       subtitle = "A ratio of 1 means there was no increase; A ratio of 1.6 means there was a 60% increase of home deaths in 2020 over the mean 2015-2019.",
       caption = paste0(c(areas_relatively_higher_number_home_deaths, home_deaths_ratio, source_nrs_text), collapse="\n"),
       fill = "Ratio"
       ) +
  NULL

@@

```{r}

>>>>

Note: the cumulative case counts have already been calculated, we just need to subset them to match the weeks. This is not trivial - the death counts cover an entire week starting from the "w/c date"; the cumulative cases for that week could be taken either at the start, the end of the week, or the start of the following week even!

However, when calculating weekly positive cases we can simply take the maximum of that week's cumulative count - that will be the cumulative count at the end of the week!

Note also that the `crude_rate_positive_per100k` represents the cumulative rate!

@@

```{r}
areas_relatively_higher_number_home_deaths_text <- "Lighter areas represent areas with relatively higher number of home deaths in 2020 compared to historical mean"
home_deaths_ratio_text_text <- "Ratio computed from total home deaths in 2020 divided by mean annual home deaths 2015-2019"

@@

```{r, fig.height=12, fig.width=8}
# TODO: work out the size of these plots
figure_map_ratio_home_deaths_2020_to_historical_la <-
  total_deaths_la %>%
  filter(
    year == 2020 & place_of_death == "Home & other non-institution"
  ) %>%
  left_join(shapefile_la, by=c("ref_area"="LAD20NM")) %>%
  ggplot(aes(geometry=geometry, fill=ratio_annual_deaths_to_historical_mean), colour="grey10") +
  geom_sf(colour="grey10") +
  scale_fill_viridis_b(limits = c(minimum_ratio_between_hb_and_la_home_deaths, maximum_ratio_between_hb_and_la_home_deaths)) +
  theme_minimal() +
  theme(axis.text = element_blank()) +
  labs(x="", y="",
       title = str_wrap("Ratio of home deaths in 2020 to historical mean 2015-2019, by Local Authority", width = 50),
       subtitle = str_wrap("A ratio of 1 means there was no increase; A ratio of 1.6 means there was a 60% increase of home deaths in 2020 over the mean 2015-2019.", width = 80),
       caption = paste0(c(areas_relatively_higher_number_home_deaths_text, home_deaths_ratio_text, source_nrs_text), collapse="\n"),
       fill = "Ratio"
       ) +
  NULL

@@

```{r, fig.height=12, fig.width=8}
# TODO: work out the size of these plots
figure_map_ratio_home_deaths_2020_to_historical_hb <-
  total_deaths_hb %>%
  filter(
    year == 2020 & place_of_death == "Home & other non-institution"
  ) %>%
  left_join(shapefile_hb_uk_wide, by=c("ref_area"="name")) %>%
  ggplot(aes(geometry=geometry, fill=ratio_annual_deaths_to_historical_mean), colour="grey10") +
  geom_sf(colour="grey10") +
  scale_fill_viridis_b(limits = c(minimum_ratio_between_hb_and_la_home_deaths, maximum_ratio_between_hb_and_la_home_deaths)) +
  theme_minimal() +
  theme(axis.text = element_blank(), aspect.ratio = 1.4) +
  labs(x="", y="",
       title = str_wrap("Ratio of home deaths in 2020 to historical mean 2015-2019, by Health Board", width=50),
       subtitle = str_wrap("A ratio of 1 means there was no increase; A ratio of 1.6 means there was a 60% increase of home deaths in 2020 over the mean 2015-2019.", width = 80),
       caption = paste0(c(areas_relatively_higher_number_home_deaths_text, home_deaths_ratio_text, source_nrs_text), collapse="\n"),
       fill = "Ratio"
       ) +
  NULL

@@

## Per capita cases of COVID-19 by region

```{r}
areas_relatively_higher_number_cases <- "Lighter areas represent areas with relatively higher number of COVID-19 cases per 100,000 population in 2020"
```

```{r, fig.height=12, fig.width=8}
figure_map_per_capita_covid_cases_2020_la <-
  weekly_covid_cases_la %>%
  filter(year == 2020 & week_number == 53) %>%
  left_join(shapefile_la, by=c("ref_area"="LAD20NM")) %>%
  ggplot(aes(geometry=geometry, fill=crude_rate_positive_per100k), colour="grey10") +
  geom_sf(colour="grey10") +
  scale_fill_viridis_b(option = "B") +
  theme_minimal() +
  theme(axis.text = element_blank()) +
  labs(x="", y="",
       title = str_wrap("Cumulative cases of COVID-19 in 2020 per 100,000 population, by Local Authority",width=50),
       subtitle = "",
       caption = paste0(c(areas_relatively_higher_number_cases, source_phs_text), collapse="\n"),
       fill = str_wrap("COVID-19 cases in 2020 per 100,000 population", width = 20)
       ) +
  NULL

walk(
  .x = c(".pdf",".svg",".png"),
  .f = ~ggsave(filename = paste0(dir_outputs,"/figure_map_per_capita_covid_cases_2020_la",.x), plot = figure_map_per_capita_covid_cases_2020_la, width = 9, height = 12, dpi = 300)
)

figure_map_per_capita_covid_cases_2020_la
```

```{r, fig.height=12, fig.width=8}
figure_map_per_capita_covid_cases_2020_hb <-
  weekly_covid_cases_hb %>%
  filter(year == 2020 & week_number == 53) %>%
  left_join(shapefile_hb_uk_wide, by=c("ref_area"="name")) %>%
  ggplot(aes(geometry=geometry, fill=crude_rate_positive_per100k), colour="grey10") +
  geom_sf(colour="grey10") +
  scale_fill_viridis_b(option = "B") +
  theme_minimal() +
  theme(axis.text = element_blank(), aspect.ratio = 1.4) +
  labs(x="", y="",
       title = str_wrap("Cumulative cases of COVID-19 in 2020 per 100,000 population, by Health Board",width=50),
       subtitle = "",
       caption = paste0(c(areas_relatively_higher_number_cases, source_phs_text), collapse="\n"),
       fill = str_wrap("COVID-19 cases in 2020 per 100,000 population", width = 20)
       ) +
  NULL

walk(
  .x = c(".pdf",".svg",".png"),
  .f = ~ggsave(filename = paste0(dir_outputs,"/figure_map_per_capita_covid_cases_2020_hb",.x), plot = figure_map_per_capita_covid_cases_2020_hb, width = 9, height = 12, dpi = 300)
)

figure_map_per_capita_covid_cases_2020_hb
```
